Trajectories

### 2022-09-16 index.qmd
Updte report

---
title: "Trajectories"
subtitle: "Lab notes"
date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
authors:
  - name: Gonzalo Garcia-Castro
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    corresponding: true
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
        address: Ramon Trias fargas 25-27
        city: Barcelona
        state: Spain
        postal-code: 08005
  - name: Daniela S. Ávila-Varela
    orcid: 0000-0002-3518-8117
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
  - name: Ignacio Castillejo
    orcid: 0000-0001-7445-0416
    affiliations:
      - name: Universidad Atónoma de Madrid
        department: Departamento de Psicología
  - name: Nuria Sebastian-Galles
    orcid: 0000-0001-6938-2498
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al. (2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
format:
    html:
        toc: true
        smooth-scroll: true
        link-external-newwindow: true
        number-sections: true
        number-offset: 0
        code-fold: true
        code-overflow: scroll
        code-line-numbers: true
        code-copy: hover
        fig-dpi: 500
        reference-location: margin
        execute-dir: "C:/Users/U155880/Documents/trajectories/"
bibliography: "manuscript/references.bib"
csl: "manuscript/apa6.csl"
---

@@

library(bayesplot)
library(brms) 
library(conflicted)
library(dplyr) 
library(forcats)
library(ggplot2) 
library(marginaleffects)
library(ggsci)
library(gt) 
library(here) 
library(janitor)
library(knitr)
library(lubridate)
library(multilex)
library(patchwork) 
library(papaja) 
library(purrr)
library(scales) 
library(stringdist)
library(stringr)
library(targets)
library(tibble)
library(tidybayes) 
library(tidyr)

@@

```{r}
#| label: items-summary
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Lexical frequencies. Mean, standard error, and 95\\% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))

@@

```{r}
#| label: items-frequency
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    ggplot() +
    aes(
        x = freq,
        fill = language,
        colour = language,
        shape = language
    ) +
    stat_slab(
        slab_colour = "white",
        slab_size = 1,
        alpha = 0.5,
    ) +
    stat_pointinterval(
        aes(y = -0.1),
        position = position_dodge(width = 0.1),
        point_size = 3
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "Frequency (Zipf)",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent,  limits = c(-0.15, 1), breaks = seq(0, 1, 0.25)) +
    theme(
        legend.position = "top",
        legend.title = element_blank()
    )
```

@@

```{r}
#| label: items-nphon
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>%
    count(language, n_phon) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(
        x = n_phon,
        y = prop,
        fill = language,
        colour = language,
        shape = language
    ) +
    geom_col(
        position = position_dodge(width = 0.95)
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "# Phonemes",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    scale_x_continuous(breaks = seq(0, 20, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(),
        legend.title = element_blank()
    )
```

@@

```{r}
#| label: items-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv_cut = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv_cut) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(x = lv_cut, y = prop) +
    geom_col(fill = ggsci::pal_d3()(4)[1]) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    labs(
        x = "Levenshtein",
        y = "Probability density",
    ) +
    theme(
        legend.position = "top"
    )
```

@@

```{r}
#| label: items-frequency-phonemes
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, freq, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: items-frequency-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(freq, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.5) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: items-nphon-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: participants-time
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants <- left_join(
    participants, 
    select(multilex_data$logs, id, time, edu_parent, dominant_language = dominance)
)

@@

```{r}
#| label: participants-age
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = pal_d3()(3)[1]) +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

@@

```{r}
#| label: participants-lp
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants %>% 
    mutate(
        age = round(age),
        id = paste(id, "(", time, ")"),
        rank = rank(doe_catalan)
    ) %>% 
    pivot_longer(
        starts_with("doe_"), 
        names_to = "language",
        values_to = "doe",
        names_prefix = "doe_",
        names_transform = list(language = str_to_sentence)
    ) %>% 
    ggplot() +
    aes(x = doe, y = reorder(id, -rank), fill = language) +
    geom_col(width = 1, position = position_fill(), colour = "white", size = 0.1) +
    geom_vline(xintercept = seq(0, 1, 0.25), colour = "grey") +
    labs(
        x = "Degree of Exposure (DoE)",
        y = "Participant",
        colour = "Language",
        fill = "Language"
    ) +
    scale_x_continuous(labels = percent) +
    theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.title = element_blank()
    )
```

@@

```{r}
#| label: participants-edu
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants %>%
    mutate(
        edu_parent = fct_explicit_na(as.factor(edu_parent)),
        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
    ) %>% 
    count(doe_catalan, edu_parent) %>% 
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(multilex_data$logs$edu_parent)
        )
    ) %>% 
    rename(total = n) %>% 
    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
    rename(Total = total) %>% 
    select(-"NA") %>% 
    gt() %>% 
    cols_label(
        doe_catalan = "DoE Catalan"
    ) %>% 
    summary_rows(
        columns = 2:7, 
        fns = list(Total = ~sum(., na.rm = TRUE)),
        formatter = fmt_number,
        decimals = 0
    ) %>% 
    fmt_missing(columns = 3:8, missing_text = "--") %>% 
    tab_spanner("Educational attainment", 3:8) %>% 
    tab_style(
        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
        locations = cells_body(columns = 9, rows = 1:10)
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

```{r}
#| label: participants-edu-plot
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants %>%
    left_join(select(multilex_data$logs, id, time, edu_parent)) %>% 
    mutate(
        doe_2 = pmin(doe_catalan, doe_spanish),
        doe_2 = cut(doe_2, seq(0, 0.5, 0.1), include.lowest = TRUE)
    ) %>% 
    count(doe_2, edu_parent) %>% 
    ggplot() +
    aes(
        x = doe_2,
        y = n,
        fill = edu_parent
    ) +
    geom_col(position = position_fill()) +
    scale_fill_manual(values = sort(pal_d3()(6), decreasing = TRUE)) +
    labs(
        x = "Exposure to second language (L2 DoE)",
        y = "Proportion of participants",
        fill = "Highest parental educational attainment"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid = element_blank()
    )
```

@@

```{r}
#| label: prior-predictions
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
nd <- expand.grid(
    n_phon_std = 0,
    freq_std = 0,
    age_std = (10:36-mean(df$age))/(sd(df$age)),
    doe_std = 0,
    lv_std = 0
)

@@

m %>% 
    ggplot() +
    aes(
        x = age_std, 
        y = .epred, 
    ) +
    facet_wrap(~.category) +
    # geom_line(
    #     aes(group = interaction(.draw)),
    #     alpha = 0.05,
    #     size = 1
    # ) +
    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
    stat_summary(fun = "mean", geom = "line", size = 0.75) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4),
        sec.axis = dup_axis(
            name = "Age (years; months)",
            labels = months_to_years(
                rescale_variable(
                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
                    mean(df$age),
                    sd(df$age)
                )
            )
        )
    ) +
    scale_fill_brewer(
        palette = "YlOrRd",
        labels = c("95%", "89%", "67%", "50%")
    ) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition\n(Linear prediction)",
        colour = "Credible interval", 
        fill = "Credible interval"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

@@

We compared the performance of these models using Bayesian leave-one-out cross-validation (LOO) using the `loo` and `loo_compare` functions of the `brms` R package (dependent of the `LOO` R package). LOO consists in computing the average likelihood of each observation after estimating the model's parameters leave that same observation out of the data set. Although the `loo` function uses a particular algorithm that speeds up the computation of this criterion (pareto-smooth importance sampling, PSIS), the size of our data set lead us to rely on the computation of the same criterion using a sampling approach via de `loo_subsample` function.

```{r}
#| label: results-loos
#| echo: true
#| message: false
#| warning: false
#| eval: false
#| fig-width: 7
model_loos %>% 
    loo_compare() %>% 
    as.data.frame() %>%   
    rownames_to_column("model") %>% 
    relocate(
        model, 
        matches("elpd_loo"),
        matches("p_loo"), 
        matches("looic"), 
        matches("diff")
    ) %>%
    gt() %>% 
    tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
    tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
    tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
    tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
    fmt_number(2:9) %>% 
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("SE"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*")
    ) %>% 
    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)"))
```

@@

```{r}
#| label: loo-diagnostics-1
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
get_loo_plot(model_loos$model_fit_1)
```

>>>>

---
title: "Trajectories"
subtitle: "Lab notes"
date: "Updated: `r format(Sys.Date(), '%Y/%m/%d')`"
authors:
  - name: Gonzalo Garcia-Castro
    orcid: 0000-0002-8553-4209
    email: gonzalo.garciadecastro@upf.edu
    corresponding: true
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
        address: Ramon Trias fargas 25-27
        city: Barcelona
        state: Spain
        postal-code: 08005
  - name: Daniela S. Ávila-Varela
    orcid: 0000-0002-3518-8117
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
  - name: Ignacio Castillejo
    orcid: 0000-0001-7445-0416
    affiliations:
      - name: Universidad Atónoma de Madrid
        department: Departamento de Psicología
  - name: Nuria Sebastian-Galles
    orcid: 0000-0001-6938-2498
    affiliations:
      - name: Universitat Pompeu Fabra
        department: Center for Brain and Cognition
abstract: "Bilinguals face the challenging task of learning words from languages with overlapping phonologies. Floccia et al. (2018) reported larger vocabulary sizes for 24-month-old bilinguals that were learning languages that shared a greater amount of cognates (e.g., English-Dutch). The mechanisms underlying this effect remain unknown. We explore two compatible scenarios. First, we test whether cognates are learnt earlier than non-cognates. This would account for the difference in vocabulary size associated to the amount of shared cognates across languages. Second, we explore the possibility that the word-forms of one language interact with those form the other language, scaffolding the acquisition of their translation equivalents when their phonologies overlap. This mechanism, in line with the parallel activation account of bilingual speech perception, would provide a plausible explanation to why cognates are acquired ealier by bilinguals. We developed an online tool to collect parental reports of receptive and productive vocabularies from children learning Catalan and/or Spanish, and present data on receptive and productive vocabulary of bilingual toddlers aged 12 to 34 months."
format:
    html:
        toc: true
        link-external-newwindow: true
        number-sections: true
        number-offset: 0
        code-fold: true
        code-overflow: scroll
        code-line-numbers: true
        code-copy: hover
        fig-dpi: 1000
        reference-location: margin
        execute-dir: "C:/Users/U155880/Documents/trajectories/"
    pdf:
        toc: true
        number-sections: true
        colorlinks: true
        fig-dpi: 1000
bibliography: "references.bib"
csl: "manuscript/apa6.csl"
---

@@

library(bayesplot)
library(brms) 
library(bayestestR)
library(conflicted)
library(dplyr) 
library(forcats)
library(ggplot2) 
library(ggsci)
library(gt) 
library(here) 
library(janitor)
library(knitr)
library(lubridate)
library(multilex)
library(patchwork) 
library(papaja) 
library(purrr)
library(scales) 
library(stringdist)
library(stringr)
library(targets)
library(tibble)
library(tidybayes) 
library(tidyr)

@@

```{r}
#| label: tbl-items-summary
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Lexical frequencies. Mean, standard error, and 95% confidence interval of lexical frequencies of items included in the Catalan and Spanish lists, reported separately for identical cognates, non-identical cognates, and non-cognates."
items <- left_join(items, select(multilex_data$pool, te, item, language, class, category))

@@

```{r}
#| label: fig-items-freq
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of lexical frequency across items"
items %>% 
    ggplot() +
    aes(
        x = freq,
        fill = language,
        colour = language,
        shape = language
    ) +
    stat_slab(
        slab_colour = "white",
        slab_size = 1,
        alpha = 0.75,
    ) +
    stat_pointinterval(
        aes(y = -0.1),
        position = position_dodge(width = 0.1),
        point_size = 3
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "Frequency (Zipf)",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent,  limits = c(-0.15, 1), breaks = seq(0, 1, 0.25)) +
    theme(
        legend.position = "top",
        legend.title = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of number of phonemes across items"
items %>%
    count(language, n_phon) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(
        x = n_phon,
        y = prop,
        fill = language,
        colour = language,
        shape = language
    ) +
    geom_col(
        position = position_dodge(width = 0.95)
    ) +
    geom_hline(yintercept = 0, colour = "black", size = 0.5) +
    labs(
        x = "# Phonemes",
        y = "Probability density",
        fill = "Language",
        shape = "Language",
        colour = "Language"
    ) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    scale_x_continuous(breaks = seq(0, 20, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(),
        legend.title = element_blank()
    )
```

@@

```{r}
#| label: fig-items-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of Levenshtein similarities across items"
items %>% 
    distinct(te, .keep_all = TRUE) %>% 
    mutate(lv_cut = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
    count(lv_cut) %>% 
    mutate(prop = n/sum(.$n)) %>% 
    ggplot() +
    aes(x = lv_cut, y = prop) +
    geom_col(fill = ggsci::pal_d3()(4)[1]) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.025)) +
    labs(
        x = "Levenshtein",
        y = "Probability density",
    ) +
    theme(
        legend.position = "top"
    )
```

@@

```{r}
#| label: fig-items-frequency-phonemes
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen number of lexical frequency and number of phonemes"
items %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, freq, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: fig-items-frequency-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen lexical frequency and phonological similarity"
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(freq, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.5) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        y = "Levenshtein similarity", 
        x = "Lexical frequency (Zipf score)\nExtracted from SUBTLEX", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_blank()
    )
```

@@

```{r}
#| label: fig-items-nphon-lv
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Association betwen number of phonemes and phonological similarity"
items %>% 
    left_join(select(pool, item, language)) %>% 
    drop_na(freq) %>% 
    ggplot(aes(n_phon, lv, colour = language, fill = language)) +
    facet_wrap(~language) +
    geom_point(size = 1, alpha = 0.25) +
    geom_smooth(method = "lm", formula = "y ~ x") +
    labs(
        x = "# Phonemes", 
        y = "Levenshtein similarity", 
        colour = "Language", 
        fill = "Language"
    ) +
    scale_y_continuous(labels = percent) +
    scale_x_continuous(breaks = seq(0, 16, 1)) +
    theme(
        legend.position = "none",
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank()
    ) 
```

@@

```{r}
#| label: fig-participants-time
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Data at which responses were collected"
participants <- left_join(
    participants, 
    select(multilex_data$logs, id, time, edu_parent, dominant_language = dominance)
)

@@

```{r}
#| label: fig-participants-age
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Distribution of participants' ages"
participants %>% 
    mutate(age = floor(age)) %>% 
    count(age) %>% 
    ggplot() +
    aes(x = age, y = n) +
    geom_col(fill = pal_d3()(3)[1]) +
    geom_text(aes(label = n), size = 3.5, vjust = -1) +
    labs(
        x = "Age (months)",
        y = "# participants"
    ) +
    scale_x_continuous(breaks = seq(0, 40, 2)) +
    scale_y_continuous(limits = c(0, 55), breaks = seq(0, 55, 10)) +
    theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank()
    )
```

@@

```{r}
#| label: fig-participants-lp
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Participant degree of exposure (DoE) to Catalan and Spanish"
participants %>% 
    mutate(
        age = round(age),
        id = paste(id, "(", time, ")"),
        rank = rank(doe_catalan)
    ) %>% 
    pivot_longer(
        starts_with("doe_"), 
        names_to = "language",
        values_to = "doe",
        names_prefix = "doe_",
        names_transform = list(language = str_to_sentence)
    ) %>% 
    ggplot() +
    aes(x = doe, y = reorder(id, -rank), fill = language) +
    geom_col(width = 1, position = position_fill(), colour = "white", size = 0.1) +
    geom_vline(xintercept = seq(0, 1, 0.25), colour = "grey") +
    labs(
        x = "Degree of Exposure (DoE)",
        y = "Participant",
        colour = "Language",
        fill = "Language"
    ) +
    scale_x_continuous(labels = percent) +
    theme(
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        axis.title = element_blank()
    )
```

@@

```{r}
#| label: tbl-participants-edu
#| tbl-cap: Participant exposure to the language (*DoE*) and socio-economic status (*SES*)
#| echo: true
#| message: false
#| warning: false
participants %>%
    mutate(
        edu_parent = fct_explicit_na(as.factor(edu_parent)),
        doe_catalan = cut(doe_catalan, breaks = seq(0, 1, 0.1), include.lowest = TRUE)
    ) %>% 
    count(doe_catalan, edu_parent) %>% 
    right_join(
        expand_grid(
            doe_catalan = levels(.$doe_catalan),
            edu_parent = unique(multilex_data$logs$edu_parent)
        )
    ) %>% 
    rename(total = n) %>% 
    pivot_wider(names_from = edu_parent, values_from = total, values_fill = 0) %>%
    mutate(total = rowSums(cbind(.[,3:8]), na.rm = TRUE)) %>% 
    rename(Total = total) %>% 
    select(-"NA") %>% 
    gt() %>% 
    cols_label(
        doe_catalan = "DoE Catalan"
    ) %>% 
    summary_rows(
        columns = 2:7, 
        fns = list(Total = ~sum(., na.rm = TRUE)),
        formatter = fmt_number,
        decimals = 0
    ) %>% 
    sub_missing(columns = 2:8, missing_text = "--") %>% 
    tab_spanner("Educational attainment", 3:8) %>% 
    tab_style(
        style = list(cell_borders(weight = px(2), sides = "left", color = "grey50")),
        locations = cells_body(columns = 9, rows = 1:10)
    ) %>% 
    tab_style(
        cell_text(style = "italic"),
        cells_column_labels(columns = 2:7)
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_labels(columns = c(1, 8))
    ) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners(spanners = "Educational attainment")
    )
```

```{r}
#| label: fig-participants-edu
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
participants %>%
    left_join(select(multilex_data$logs, id, time, edu_parent)) %>% 
    mutate(
        doe_2 = pmin(doe_catalan, doe_spanish),
        doe_2 = cut(doe_2, seq(0, 0.5, 0.1), include.lowest = TRUE)
    ) %>% 
    count(doe_2, edu_parent) %>% 
    ggplot() +
    aes(
        x = doe_2,
        y = n,
        fill = edu_parent
    ) +
    geom_col(position = position_fill()) +
    scale_fill_manual(values = sort(pal_d3()(6), decreasing = TRUE)) +
    labs(
        x = "Exposure to second language (L2 DoE)",
        y = "Proportion of participants",
        fill = "Highest parental educational attainment"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        panel.grid = element_blank()
    )
```

@@

```{r}
#| label: fig-prior-predictions
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Expected prior predictions."
nd <- expand.grid(
    n_phon_std = 0,
    freq_std = 0,
    age_std = (10:36-mean(df$age))/(sd(df$age)),
    doe_std = 0,
    lv_std = 0
)

@@

m %>% 
    ggplot() +
    aes(
        x = age_std, 
        y = .epred, 
    ) +
    facet_wrap(~.category) +
    # geom_line(
    #     aes(group = interaction(.draw)),
    #     alpha = 0.05,
    #     size = 1
    # ) +
    stat_lineribbon(size = 1, colour = NA, .width = c(0.95, 0.89, 0.67, 0.5)) +
    stat_summary(fun = "mean", geom = "line", size = 0.75) +
    scale_x_continuous(
        breaks = (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
        labels = seq(10, 36, 4),
        sec.axis = dup_axis(
            name = "Age (years; months)",
            labels = months_to_years(
                rescale_variable(
                    (seq(10, 36, 4)-mean(df$age))/(sd(df$age)),
                    mean(df$age),
                    sd(df$age)
                )
            )
        )
    ) +
    scale_fill_brewer(
        palette = "Reds",
        labels = c("95%", "89%", "67%", "50%")
    ) +
    labs(
        x = "Age (months)", 
        y = "Probability of acquisition\n(Linear prediction)",
        colour = "Credible interval", 
        fill = "Credible interval"
    ) +
    scale_y_continuous(labels = percent, limits = c(0, 1)) +
    theme(
        legend.position = "top",
        axis.ticks.x = element_line(colour = "black"),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.border = element_rect(fill = NA, colour = "black", size = 0.75),
        strip.background = element_rect(fill = "grey", colour = NA),
        plot.background = element_rect(fill = "white", colour = NA)
    )
```

@@

We compared the predictive performance of the models using Bayesian leave-one-out cross-validation (LOO) via the `loo` R package. LOO consists in computing the average likelihood of each observation after estimating the model's parameters leaving that same observation out of the data set. Due to computational constraints, we performed LOO based on a random subsample of 750 samples (out of `r format(dim(model_fit_3$fit)[1], big.mark = ",")`) of the posterior distribution of the fixed effects of the model as:

```{r}
#| label: results-loos
#| echo: true
#| message: false
#| warning: false
#| eval: false
#  loo(log_lik(model_fit_3, ndraws = 750, re_formula = NA))
```

@tbl-results-loos shows the outputs of the *LOO-CV*. The extended model (`model_fit_3`), which included the three-way interaction between $Age$, $DoE$, and $Levenshtein$ showed the best predictive performance, with a difference in *ELPD* compared to the other models several times larger than the standard error of such difference. This indicates that, given the data, the predictions of the extended model are confidently more accurate than those of the other models.

```{r}
#| label: tbl-results-loos
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| tbl-cap: "Bayesian leave-one-out cross validation (LOO-CV)"
model_loos %>% 
    loo_compare() %>% 
    as.data.frame() %>%   
    rownames_to_column("model") %>% 
    relocate(
        model, 
        matches("elpd_loo"),
        matches("p_loo"), 
        matches("looic"), 
        matches("diff")
    ) %>%
    gt() %>% 
    tab_header(
        title = "",
        subtitle = md("The predictive accuracy of the models was compared using leave-one-out cross-validation. The likelihood of each data point given a model fitted without including that same observation was estimated. The resulting likelihoods of each model were summarised as its expected log-predicted density (*ELPD*), which penalises the complexity of each model. *ELPD* values closer to zero indicate better predictive acuracy. We report the *ELPD*, effective number of parameters (*p*), and information criterion (*IC*) of each model along their associated standard errors (*SE*). The last column indicates the difference in *ELPD* between each model and the model with the best predictive accuracy (`model_fit_3`)")
    ) %>% 
    tab_spanner(md("LOO<sub>ELPD</sub>"), matches("elpd_loo")) %>% 
    tab_spanner(md("LOO<sub>p</sub>"), matches("p_loo")) %>% 
    tab_spanner(md("LOO<sub>IC</sub>"), matches("looic")) %>% 
    tab_spanner(md("LOO<sub>diff</sub>"), matches("diff")) %>% 
    fmt_number(2:9) %>% 
    cols_label(
        model = "Model",
        # elpd
        elpd_loo = md("*ELPD*"),
        se_elpd_loo = md("*SE*"),
        # p
        p_loo = md("*p*"),
        se_p_loo = md("*SE*"),
        # looic
        looic = md("*LOO-IC*"),
        se_looic = md("*SE*"),
        # diff
        elpd_diff = md("*diff*"),
        se_diff = md("*SE*")
    ) %>% 
    tab_source_note(md("Pareto-*k* estimates of all models were acceptable (*k* < 0.5)")) %>% 
    tab_style(
        cell_text(weight = "bold"),
        cells_column_spanners()
    ) %>% 
    tab_style(
        cell_text(align = "left"),
        cells_title(groups = "subtitle")
    )
```

@@

```{r}
#| label: fig-loo-diagnostics-1
#| echo: true
#| message: false
#| warning: false
#| fig-width: 7
#| fig-cap: "Bayesian leave-out-out cross-validation diagnostics of Model 1"
get_loo_plot(model_loos$model_fit_1)
```
